Robustly decorrelating errors with mixed quantum gates

### 2020-03-05 decorrelation/mixed_unitaries.tex
changes to sparisty section

\subsubsection{Sparsity Constraints}
\label{sec:sparsity}
\noindent
As a practical consideration, we would also like to regularize our objective function to promote sparse weightings. Control electronics often have a limited amount of waveform memory, and thus it is important to be able to construct MQGs with as few constituent gates as possible. In many machine learning contexts, lasso regularization \cite{tibshirani1996regression} can be used to enforce sparsity in solutions. Rather than penalizing the $\ell_0$ norm of the solution, which is not convex, lasso regularization penalizes the $\ell_1$, which is the convex relaxation of the $\ell_0$ norm. However this is insufficient for our purposes, as we already require $\weight$ to be a valid probability distribution, which constrains the $\ell_1$ norm to be equal to $1$. Conveniently, the problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends Eq.~\eqref{eq:minimization}:
\begin{equation}\label{eq:minimization_regularization}
\begin{split}
&\underset{m\in[1..N]}{\textbf{minimize:}}\\
&\ \ \bunderbrace{\textbf{\hspace{.2cm}minimize:\hspace{.2cm}} }{\hspace{.1cm}\weight_i\geq0, |\weight|_1=1, t\geq0} ||\mathbf{L}\cdot\vec\weight||_2 + t\\
&\ \ \hspace{.2cm}\textbf{subject to: } \weight_m > \frac{\lambda}{t}.
\end{split}
\end{equation}
Here, $\lambda > 0$ is a tunable parameter that can be used to control the degree of sparsity in the solution. Increasing $\lambda$ will increase the sparsity of the solution, but will generally decrease the likelihood that the solution will be generator-exact.

@@

\begin{figure}
\includegraphics[scale=.5]{sparisity-versus-parameter.png}
\caption{Sparsity of MQGs as a function of the sparisty parameter $\lambda$, with higher weight controls omitted to emphasize the low-weight distribution. Numerical values of $\lambda_i$ are: $\lambda_0=0$, $\lambda_1=6.25$e-05, $\lambda_2=0.0001$, $\lambda_3=0.00025$ and $\lambda_4=0.0005$. As the parameter is increased from zero, the number of controls with low probability increases. For $\lambda_4$ virtually all of the 750 controls have vanishingly small probability, allowing for the restriction of the MQG to a small number of controls.}
\label{fig:sparsity}
\end{figure}
To generate the initial controls, we use the GRAPE algorithm\cite{Khaneja2005} with N=25 steps and total evolution time of $\pi$ to generate 750 candidate controls. However, by imposing the sparisty constraint discussed in Section \ref{sec:sparsity}, we found MQGs consisting of just 10 controls. Fig.~\ref{fig:sparsity} shows that as $\lambda$ (Eq. \eqref{eq:minimization_regularization}) increases, the support of the MQGs become more peaked, with most of their probability mass concentrating on a smaller number of controls. In our implementation of the GRAPE algorithm, we use the performance function presented in \cite{Khaneja2005}, and average over different values of $\delta$ and $\epsilon$ using Gaussian quadrature when computing the gradient, so that we find controls that are naturally robust. The standard deviations considered for all parameters in our numerical experiments were fixed to $\sigma=.001$. Finally, we assume that the errors on $\sigma_x$ and $\sigma_y$ are perfectly correlated, as in our experimental implementation. We note that advanced quantum control protocols may provide an even more principled approach to control synthesis. DMORPH \cite{dominy2008exploring}, for example, explores continuous families of controls on fidelity level sets, thereby enabling further optimizations against secondary criteria, such as the duration of the control pulse or robustness to drift.

>>>>

\subsubsection{Sparsity Constraints}
\label{sec:sparsity}
\noindent
As a practical consideration, we would also like to regularize our objective function to promote sparse weightings, resulting in MQGs with only a small number of gate implementations. Control electronics often have a limited amount of waveform memory, and thus it is important to be able to construct MQGs with as few constituent gates as possible. In many machine learning contexts, a technique called lasso regularization \cite{tibshirani1996regression} can be used to enforce this sparsity in solutions.

Rather than penalizing the $\ell_0$ norm of the solution, which counts the number of nonzero entries in a vector, lasso regularization penalizes the $\ell_1$ norm, which is a convex relaxation of the $\ell_0$ norm. Because the extreme points of the unit $\mathcal{L}_1$ sphere lie on the unit Cartesian vectors, solutions to the lasso regularized problem are guaranteed to be sparse. However this is insufficient for our purposes, as we already require $\weight$ to be a valid probability distribution, which constrains the $\ell_1$ norm to be equal to $1$.

The problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends Eq.~\eqref{eq:minimization}:
\begin{equation}\label{eq:minimization_regularization}
\begin{split}
&\underset{m\in[1..N]}{\textbf{minimize:}}\\
&\ \ \bunderbrace{\textbf{\hspace{.2cm}minimize:\hspace{.2cm}} }{\hspace{.1cm}\weight_i\geq0, |\weight|_1=1, t\geq0} ||\mathbf{L}\cdot\vec\weight||_2 + t\\
&\ \ \hspace{.2cm}\textbf{subject to: } \weight_m > \frac{\lambda}{t}.
\end{split}
\end{equation}
Here, $\lambda \geq 0$ is a tunable parameter that can be used to control the degree of sparsity in the solution. If $\lambda$ is set to zero, the inner program reduces to Eq.~\eqref{eq:minimization}, and the outer minimization is redundant. If $\lambda$ is very large each inner minimization problem reduces to a MQG with a single control selected. Therefore the outer minimization simply selects the control whose corresponding error generator has smallest norm. While this gives a maximally sparse solution, it doesn't allow for any cancellation of coherent errors. In particular, increasing $\lambda$ will increase the sparsity of the solution, but will generally decrease the likelihood that the solution will be generator-exact.

@@

\begin{figure}
\includegraphics[scale=.5]{sparisity-versus-parameter.png}
\caption{Sparsity of MQGs as a function of the sparisty parameter $\lambda$, with higher weight controls omitted to emphasize the low-weight distribution. Numerical values of $\lambda_i$ are: $\lambda_0=0$, $\lambda_1=6.25$e-05, $\lambda_2=0.0001$, $\lambda_3=0.00025$ and $\lambda_4=0.0005$. As the parameter is increased from zero, the number of controls with low probability increases. For $\lambda_4$ virtually all of the 750 controls have vanishingly small probability, allowing for the restriction of the MQG to a small number of controls.}
\label{fig:sparsity}
\end{figure}
To generate the initial controls, we use the GRAPE algorithm\cite{Khaneja2005} with N=25 steps and total evolution time of $\pi$ to generate 750 candidate controls. However, by imposing the sparsity constraint discussed in Section \ref{sec:sparsity}, we found MQGs consisting of just 10 controls. Fig.~\ref{fig:sparsity} shows that as $\lambda$ (Eq. \eqref{eq:minimization_regularization}) increases, the support of the MQGs become more peaked, with most of their probability mass concentrating on a smaller number of controls. In our implementation of the GRAPE algorithm, we use the performance function presented in \cite{Khaneja2005}, and average over different values of $\delta$ and $\epsilon$ using Gaussian quadrature when computing the gradient, so that we find controls that are naturally robust. The standard deviations considered for all parameters in our numerical experiments were fixed to $\sigma=.001$. Finally, we assume that the errors on $\sigma_x$ and $\sigma_y$ are perfectly correlated, as in our experimental implementation. We note that advanced quantum control protocols may provide an even more principled approach to control synthesis. DMORPH \cite{dominy2008exploring}, for example, explores continuous families of controls on fidelity level sets, thereby enabling further optimizations against secondary criteria, such as the duration of the control pulse or robustness to drift.
